The Concurrent Calculi Formalisation Benchmark

### 2024-02-21 short-paper/short-paper.tex
Clean up the beginning of the coinduction section.

\subsection{Coinduction}
Coinduction is fundamental to establish properties of processes that
may never terminate, which is a typical situation.
Common uses for coinduction are to define strong and weak versions of
bisimulation, barbed notions of bisimulation, and prove trace equivalence.
On the other hand, coinduction in proof assistants
is notoriously difficult and its representation
is non-uniform across developments.
E.g., \cite{Castro-Perez2021} notes that the most challenging
part of their mechanisation of multiparty session types
was the convoluted definition of infinite behaviours;
\cite{Pohjola2022} points out that the main difficulty
of the proofs concerning bisimulation lies
in dealing with existential quantifiers; and
\cite{Bengtson2016} wonders whether parametrised
coinduction as developed in Coq \cite{Hur2013} would
have helped their Isabelle mechanisation.

We base our challenge around bisimulation: a fundamental coinductively
established property for equating the
observed behaviour of potentially infinite processes. 
In this challenge, we study the notion of strong barbed bisimulation
(and bisimilarity, as in \cite{picalcbook})
over a fragment of the $\pi$-calculus and,
in particular, we focus on obtaining,
from strong barbed bisimilarity,
a congruence with respect to
parallel composition and substitution.

\alb{We need more details about what the challenge is}
\alb{I kinda object to ``seamlessly''. In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky. I guess Isabelle is better}

Our coinduction challenge concerns strong barbed bisimulation and
congruence.% \alb{we motivate why strong vs weak, but not why barbed}
% \lore{Is it not enough to quote the Sangiorgi/Walker book
% and go by authority? :)}
Though weak barbed congruence is a more common behavioural equivalence
w.r.t.\ the \(\pi\)-calculus, here we prefer strong equivalences to avoid the
need to abstract over the number of internal transitions, thus
simplifying the theory.  We exclude delegation from the coinduction
challenge since it is orthogonal to our primary aim of exploring
coinductive proof techniques.

The introduction of strong barbed bisimulation is one of the first
steps when studying the (observable) behaviour of process calculi,
both in classical textbooks (e.g., \cite{picalcbook}) and in existing
mechanisations, where we see a variety of different formalisation
choices, sometimes ad-hoc for the specific result or development.

In fact, coinduction is widely supported in proof assistants.
Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
Newer entrants to the field like the Lean proof assistant
are working on adding a notion of coinduction to their system~\cite{Avigad2019}%,Keizer2023}.
%
As with most features, proof assistants support
subtly different formalisms that enable different techniques and motivate this challenge.
Some systems provide more than one abstraction to use coinduction.
A good example of this are the parameterised coinduction~\cite{Hur2013}
and interaction tree~\cite{Xia2019} libraries for Coq.

\lore{Should we mention that Isabelle offers
  the command \texttt{coinduction\_upto}? Is it related?}
\todo{Explain the the point of the congruence theorem is that it should be ``easy'' once you have the other one}

>>>>

\subsection{Coinduction}
Process calculi often include constructs that allow processes to have infinite behaviours.
Coinduction is a fundamental technique for defining and reasoning about infinite objects, and it can thus be used to reason about these behaviours.
While many proof assistants support coinductive techniques, using them is notoriously difficult, and representations and proof techniques vary widely.
The crux of our challenge is the effective use of coinductive up-to techniques.
The intention is that it should be easy to use the bisimulation relation and its properties, but this may be difficult with some mechanisation approaches.

The goal of our challenge on coinductive reasoning is proving that strong barbed bisimilarity can be turned into a congruence by making it sensitive to substitution and parallel composition.
The setting for our challenge is an untyped \(\pi\)-calculus with replication of processes enabling infinite behaviours.
We do not include delegation in the calculus since it is orthogonal to our primary aim of exploring coinductive proof techniques.
We base our definition of bisimilarity on a labelled transition system semantics and an observability predicate describing the communication steps available to a process.
The introduction of strong barbed bisimulation is one of the first steps when studying the observable behaviour of process calculi, both in classical textbooks (e.g., \cite{picalcbook}) and in existing mechanisations.
Though weak barbed congruence is a more common behavioural equivalence, we prefer strong equivalences to simplify the theory by avoiding the need to abstract over the number of internal transitions in a trace.

Coinduction is widely supported in proof assistants, but they support subtly different formalisms that enable different techniques, and some systems provide more than one abstraction to use coinduction.
We thus see a variety of different formalisation choices for bisimulations, sometimes ad-hoc for the specific result.

\todo{Reorganize the rest of this section to be more coherent}
Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
Newer entrants to the field like the Lean proof assistant
are working on adding a notion of coinduction to their system~\cite{Avigad2019}.\alb{I kinda object to ``seamlessly''. In Coq it's a mess, in Abella very very limited, sized types in Agda are shaky. I guess Isabelle is better}
%
A good example of this are the parameterised coinduction~\cite{Hur2013}
and interaction tree~\cite{Xia2019} libraries for Coq.

E.g., \cite{Castro-Perez2021} notes that the most challenging
part of their mechanisation of multiparty session types
was the convoluted definition of infinite behaviours;
\cite{Pohjola2022} points out that the main difficulty
of the proofs concerning bisimulation lies
in dealing with existential quantifiers; and
\cite{Bengtson2016} wonders whether parametrised
coinduction as developed in Coq \cite{Hur2013} would
have helped their Isabelle mechanisation.

\lore{Should we mention that Isabelle offers
  the command \texttt{coinduction\_upto}? Is it related?}
